effects_rt_acc

### 2024-12-06 effects_rt_acc.Rmd
added a few new plots and tables for supps

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "brms", "tidybayes", 
         "ggdist", "future", "bayesplot", "parallel", "here", "tinytable",
         "webshot2")

@@

p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = fct_rev(term), 
                                       fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(alpha=0.7) +
  labs(title = "fixed effects of condition",
       x = "Log(RT)", y = NULL) +
  scale_fill_brewer(palette = "Dark2")
p_tidy_fixed

ggsave ("figures/manifest/b1_fixef.jpeg")
```

@@

ggsave ("figures/manifest/b1_varying_pid.jpeg")
```

@@

ggsave ("figures/manifest/b1_varyvals.jpeg")
```

@@

```{r}
post <- as_draws_df(ab1) %>% #
  select(-lp__) %>% 
  as_tibble()

post_summary <- posterior_summary(ab1, robust = TRUE)

# varying effects by participant 
# I take this to be 18 pts x 4 terms (intercept, 3 levels of group) = 72 in total
vary_vals <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(vary_vals)
```

@@

```{r}
## fixed effects of interest, excluding basic intercepts 
## Is this the right way to label the interaction?
tidy_fixed <- post %>% 
  select(starts_with("b_")) %>% # select and rename in simpler labels.
  pivot_longer(everything()) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(name, "name"), "name",
                if_else(str_detect(name, "tie"), "tie",
                if_else(str_detect(name, "both"), "both", "intercept"))),
         term = factor(term, levels=unique(term)))
head(tidy_fixed)
tail(tidy_fixed)

# remove intercepts as they are on a different scale and make it difficult to plot
tidy_fixed <- tidy_fixed %>%
  filter(!term %in% c("intercept"))
head(tidy_fixed)

p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = fct_rev(term), 
                                       fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(alpha=0.7) +
  labs(title = "fixed effects of condition",
       x = "Accuracy", y = NULL) +
  scale_fill_brewer(palette = "Dark2")
p_tidy_fixed

ggsave ("figures/manifest/ab1_fixef.jpeg")
```

@@

```{r}
## random effects of interest, including item, as well as participant and effects across pts
tidy_varying <- post %>% 
  select(contains("sd_")) %>% 
  pivot_longer(everything()) %>% 
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(name, "name"), "name",
                if_else(str_detect(name, "tie"), "tie",
                if_else(str_detect(name, "both"), "both", "intercept"))),
         term = factor(term, levels=c("intercept",
                                      "name", "tie", "both")))
head(tidy_varying)
tail(tidy_varying)
str(tidy_varying)

p_tidy_varying_pid <- ggplot(tidy_varying, aes(x = value, 
                                               y = fct_rev(term))) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(alpha = 0.7) +
  labs(title = "SD plot for varying participant effects",
       x = NULL, y = NULL) +
  coord_cartesian(xlim =c(0, 1))+
  scale_x_continuous(breaks=seq(0,1,0.1))
p_tidy_varying_pid

ggsave ("figures/manifest/ab1_varying_pid.jpeg")
```

@@

# participant level
vary_pt <- post_summary %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_PID"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(term = if_else(str_detect(param, "name"), "name",
                if_else(str_detect(param, "tie"), "tie",
                if_else(str_detect(param, "both"), "both", "intercept"))),
         term = factor(term, levels=c("intercept",
                                      "name", "tie", "both"))) %>%  
  arrange(term, Estimate) %>% 
  mutate(rank = rep(1:18, times=4),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(vary_pt)
```

@@

ggsave ("figures/manifest/ab1_varyvals.jpeg")
```

@@

# 3. plot ppreds for RT and Accuracy together #

@@

>>>>

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "brms", "tidybayes", 
         "ggdist", "future", "bayesplot", "parallel", "here", "tinytable",
         "webshot2", "rstan")

@@

p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = fct_rev(term), 
                                       fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(alpha=0.7) +
  labs(#title = "fixed effects of condition",
       x = "Log(RT)", y = NULL) +
  scale_fill_brewer(palette = "Dark2")
p_tidy_fixed

# ggsave ("figures/manifest/b1_fixef.jpeg",
#         width = 8, height = 6, dpi = 800)
```

### make a table of fixed effects ###

```{r}
b1_fixef_tab <- tidy_fixed %>%
  select(term, value) %>% 
  group_by(term) %>% 
  median_qi(value) %>% 
  mutate(dv = "reaction time") %>% 
  select(dv, term, value, .lower, .upper) 
b1_fixef_tab
```

@@

ggsave ("figures/manifest/b1_varying_pid.jpeg",
        width = 8, height = 6, dpi = 800)
```

@@

ggsave ("figures/manifest/b1_varyvals.jpeg",
        width = 8, height = 6, dpi = 800)
```

@@

```{r}
ab1_post <- as_draws_df(ab1) %>% #
  select(-lp__) %>% 
  as_tibble()

ab1_post_summary <- posterior_summary(ab1, robust = TRUE)

# all fixed effects
ab1_post_qi_b <- ab1_post %>%
  select(starts_with("b_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(ab1_post_qi_b)

# average varying effects - SD
ab1_post_qi_sd <- ab1_post %>%
  select(starts_with("sd_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(ab1_post_qi_sd)

# varying effects by participant 
# I take this to be 18 pts x 4 terms (intercept, 3 levels of group) = 72 in total
ab1_vary_vals <- ab1_post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(ab1_vary_vals)
```

@@

```{r}
## fixed effects of interest, excluding basic intercepts 
## Is this the right way to label the interaction?
ab1_tidy_fixed <- ab1_post %>% 
  select(starts_with("b_")) %>% # select and rename in simpler labels.
  pivot_longer(everything()) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(name, "name"), "name",
                if_else(str_detect(name, "tie"), "tie",
                if_else(str_detect(name, "both"), "both", "intercept"))),
         term = factor(term, levels=unique(term)))
head(ab1_tidy_fixed)
tail(ab1_tidy_fixed)

## check labels
check.labels <- ab1_tidy_fixed %>% 
  group_by(key, term) %>% 
  distinct(key, term)
check.labels

# remove intercepts as they are on a different scale and make it difficult to plot
ab1_tidy_fixed <- ab1_tidy_fixed %>%
  filter(!term %in% c("intercept"))
head(ab1_tidy_fixed)

check.labels <- ab1_tidy_fixed %>% 
  group_by(key, term) %>% 
  distinct(key, term)
check.labels

p_tidy_fixed_ab1 <- ggplot(ab1_tidy_fixed, aes(x = value, y = fct_rev(term), 
                                       fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(alpha=0.7) +
  labs(#title = "fixed effects of condition",
       x = "Accuracy", y = NULL) +
  scale_fill_brewer(palette = "Dark2")
p_tidy_fixed_ab1

# ggsave ("figures/manifest/ab1_fixef.jpeg",
#         width = 8, height = 6, dpi = 800)
```

### make a table of fixed effects ###

```{r}
ab1_fixef_tab <- ab1_tidy_fixed %>%
  select(term, value) %>% 
  group_by(term) %>% 
  median_qi(value) %>% 
  mutate(dv = "accuracy") %>% 
  select(dv, term, value, .lower, .upper) 
ab1_fixef_tab
```

@@

```{r}
## random effects of interest, including item, as well as participant and effects across pts
ab1_tidy_varying <- ab1_post %>% 
  select(contains("sd_")) %>% 
  pivot_longer(everything()) %>% 
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(name, "name"), "name",
                if_else(str_detect(name, "tie"), "tie",
                if_else(str_detect(name, "both"), "both", "intercept"))),
         term = factor(term, levels=c("intercept",
                                      "name", "tie", "both")))
head(ab1_tidy_varying)
tail(ab1_tidy_varying)
str(ab1_tidy_varying)

check.varying <- ab1_tidy_varying %>% 
  group_by(key, term) %>% 
  distinct(key, term)
check.varying

p_tidy_varying_pid_ab1 <- ggplot(ab1_tidy_varying, aes(x = value, 
                                               y = fct_rev(term))) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(alpha = 0.7) +
  labs(title = "SD plot for varying participant effects",
       x = NULL, y = NULL) +
  coord_cartesian(xlim =c(0, 1))+
  scale_x_continuous(breaks=seq(0,1,0.1))
p_tidy_varying_pid_ab1

ggsave ("figures/manifest/ab1_varying_pid.jpeg",
        width = 8, height = 6, dpi = 800)
```

@@

# participant level
ab1_vary_pt <- ab1_post_summary %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_PID"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(term = if_else(str_detect(param, "name"), "name",
                if_else(str_detect(param, "tie"), "tie",
                if_else(str_detect(param, "both"), "both", "intercept"))),
         term = factor(term, levels=c("intercept",
                                      "name", "tie", "both"))) %>%  
  arrange(term, Estimate) %>% 
  mutate(rank = rep(1:18, times=4),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(ab1_vary_pt)
```

@@

```{r}
p_tidy_varyvals_ab1 <- ggplot(ab1_vary_pt, aes(x=rank, y=Estimate,
                                           colour=outlier))+
  geom_pointrange(aes(ymin=Q2.5,ymax=Q97.5),fatten = 0.5)+
  geom_hline(yintercept=0, colour = "red")+
  scale_colour_manual(values=c("darkgrey","black")) +
  labs(title = "Varying effects by participant",
       x = "participant id")+
  facet_wrap(~term) +
  theme(legend.position = "none")
p_tidy_varyvals_ab1

ggsave ("figures/manifest/ab1_varyvals.jpeg",
        width = 8, height = 6, dpi = 800)
```

@@

# 3. make plots for RT and Accuracy together #

fixed effects

```{r}
p_fixef <- (p_tidy_fixed_ab1 | p_tidy_fixed) +
  plot_annotation(title = "Fixed effects of condition for accuracy and RT",
                  tag_levels = 'A') +
  plot_layout(guides = 'collect',
              axes = 'collect')
p_fixef

ggsave("figures/manifest/fixef.jpeg",
       width = 10, height = 6, dpi = 800)
```

posterior predictions

@@

# 4. make a table of fixed effects for rt and accuracy #

```{r}
fix_tab <- bind_rows(ab1_fixef_tab, b1_fixef_tab) %>% 
   mutate(across(where(is.double), \(x) round(x, 2)))
fix_tab

write_csv(fix_tab, "tables/manifest/fixef.csv")
```
